A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-06-01 main.tex
Wrote on Kidinding and Well-formed Kinding Rules.

% run
The main purpose of transition abstractions and transition applications are to represent {\bf{run}} operator.
In multistage calculus, {\bf{run}} is a very important operator.
It changes quoted code to the original code.
{\bf{run}} is realized with application of the empty transition $\epsilon$.
For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
because $\TB_\epsilon$ is disappeared.

@@

% omitting residualization

@@

\subsection{Dependent Types}

@@

Next, we develop \LMD by extending \LTP with \LLF-like dependent types.

@@

\begin{figure}
	\begin{center}
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \G,x:\tau@A \V M \E N : \rho @A}{\G\V\lambda x:\tau.M \E \lambda x:\sigma.N : (\Pi x:\tau.\rho)@A}{\QAbs} \\[2mm]
		\infrule{\G\V M \E L : (\Pi x:\sigma.\tau)@A \andalso \G\V N \E O : \sigma@A}{\G\V M\ N \E L\ O : \tau[x \mapsto N]@A}{\QApp} \\[2mm]
		\infrule{\G\V M \E N : \tau@{A\alpha}}{\G\V \TB_\alpha M \E \TB_\alpha N : \TW_\alpha \tau@A}{\QTB} \andalso
		\infrule{\G\V M \E N : \TW_\alpha \tau@A}{\G\V \TBL_\alpha M \E \TBL_\alpha N : \tau@{A\alpha}}{\QTBL} \\[2mm]
		\infrule{\G\V M\E N : \tau@A \andalso \alpha \notin \FTV(\G)\cup\FTV(A)}{\G\V \Lambda\alpha.M \E \Lambda\alpha.N : \forall\alpha.\tau@A}{\QGen} \\[2mm]
		\infrule{\G\V M \E N:\forall\alpha.\tau@A}{\G\V M\ \epsilon \E N\ \epsilon : \tau[\alpha \mapsto \epsilon]@A}{\QIns}\andalso
		\infrule{\G\V M \E N : \tau @A}{\G\V\%_\alpha M \E \%_\alpha N : \tau@{A\alpha}}{\QCsp} \\[2mm]
		\infrule{\G\V M:\tau@A}{\G\V M\E M : \tau@A}{\QRefl} \andalso
		\infrule{\G\V M\E N : \tau@A}{\G\V N\E M : \tau@A}{\QSym} \\[2mm]
		\infrule{\G\V M\E N : \tau@A \andalso \G\V N\E L : \tau@A}{\G\V M\E L : \tau@A}{\QTrans} \\[2mm]
		\infrule{\G,x:\sigma@A\V M:\tau@A \andalso \G\V N:\sigma@A}{\G\V(\lambda x:\sigma.M)\ N\E M[x\mapsto N] : \tau[x \mapsto N]@A}{\QBeta} \\[2mm]
		% \infrule{\G\V M:(\Pi x:\sigma.\tau)@A \andalso x\notin \text{FV}(M)}{\G\V(\lambda x:\sigma.M\ x)\E M: (\Pi x:\sigma.\tau)@A}{\QEta} \\[2mm]
		\infrule{\G\V M \E N : \tau@A}{\G\V \TBL_\alpha(\TB_\alpha M) \E N : \tau @A}{\QTBLTB} \\[2mm]
		\infrule{\G\V (\Lambda\alpha.M) : \forall\alpha.\tau@A}{\G\V (\Lambda\alpha.M)\ \epsilon \E M[\alpha \mapsto \epsilon] : \tau[\alpha \mapsto \epsilon]@A}{\QLambda} \\[2mm]
		\infrule{\G\V M:\tau@{A\alpha} \andalso \G\V M:\tau@A}{\G\V\%_\alpha M \E M : \tau@{A\alpha}}{\QPercent}
		\caption{Tem Equivalence Rules}
		\label{fig:term-equivalence-rules}
	\end{center}
\end{figure}

>>>>

% Run
The main purpose of transition abstractions and transition applications are to represent {\bf{run}} operator.
In multistage calculus, {\bf{run}} is a very important operator.
It changes quoted code to the original code.
{\bf{run}} is realized with application of the empty transition $\epsilon$.
For example, $\Lambda\alpha.(\TB_\alpha (\lambda x:\I.x))\ \epsilon$ becomes $\lambda x:\I.x$
because $\TB_\epsilon$ is disappeared.

@@

% Omitting Residualization

@@

\subsection{\LLF}

@@

Next, we develop \LMD by extending \LTP with \LLF-like dependent types.
From here, we use the word of stage instead of transition 
because \red{we develop a multistage calculus not a multitransition calculus.}

% Kidinding and Well-formed Kinding Rules

In order to extend \LMD with dependent types, we need kinding rules and well-formed kinding rules.
It was lucky that almost all rules are determined easily.
This is because multistage calculus and dependent types are almost \red{orthogonal}.
Therefore, we can get kinding rules and well-formed kinding rules of \LMD just by 
attaching stage anotations to ones of \LLF.
For example, \KAbs' is a kinding rule for a dependent type in \LLF and \KAbs\ is a corresponding one.
\begin{center}
	\infrule{\G\vdash \tau :: * \andalso \G,x:\tau@A\vdash \sigma::J}{\G\vdash(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)}{\KAbs'} \\[2mm]
	\infrule{\G\V \tau :: *@A \andalso \G,x:\tau@A\V \sigma::J@A}{\G\V(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J)@A}{\KAbs} \\[2mm]
\end{center}

% Constants and Base Types

The way of handling of constants and base types is diffrence between \LMD and \LTP.
We adopt signature $\Sigma$ to handle constants and base types.
This is because signature simplify kinding rules relating to type variables.

% Equivalence Rules

% Q-% rule

@@

\begin{figure}
	\begin{center}
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \G,x:\tau@A \V M \E N : \rho @A}{\G\V\lambda x:\tau.M \E \lambda x:\sigma.N : (\Pi x:\tau.\rho)@A}{\QAbs} \\[2mm]
		\infrule{\G\V M \E L : (\Pi x:\sigma.\tau)@A \andalso \G\V N \E O : \sigma@A}{\G\V M\ N \E L\ O : \tau[x \mapsto N]@A}{\QApp} \\[2mm]
		\infrule{\G\V M \E N : \tau@{A\alpha}}{\G\V \TB_\alpha M \E \TB_\alpha N : \TW_\alpha \tau@A}{\QTB} \andalso
		\infrule{\G\V M \E N : \TW_\alpha \tau@A}{\G\V \TBL_\alpha M \E \TBL_\alpha N : \tau@{A\alpha}}{\QTBL} \\[2mm]
		\infrule{\G\V M\E N : \tau@A \andalso \alpha \notin \FTV(\G)\cup\FTV(A)}{\G\V \Lambda\alpha.M \E \Lambda\alpha.N : \forall\alpha.\tau@A}{\QGen} \\[2mm]
		\infrule{\G\V M \E N:\forall\alpha.\tau@A}{\G\V M\ \epsilon \E N\ \epsilon : \tau[\alpha \mapsto \epsilon]@A}{\QIns}\andalso
		\infrule{\G\V M \E N : \tau @A}{\G\V\%_\alpha M \E \%_\alpha N : \tau@{A\alpha}}{\QCsp} \\[2mm]
		\infrule{\G\V M:\tau@A}{\G\V M\E M : \tau@A}{\QRefl} \andalso
		\infrule{\G\V M\E N : \tau@A}{\G\V N\E M : \tau@A}{\QSym} \\[2mm]
		\infrule{\G\V M\E N : \tau@A \andalso \G\V N\E L : \tau@A}{\G\V M\E L : \tau@A}{\QTrans} \\[2mm]
		\infrule{\G,x:\sigma@A\V M:\tau@A \andalso \G\V N:\sigma@A}{\G\V(\lambda x:\sigma.M)\ N\E M[x\mapsto N] : \tau[x \mapsto N]@A}{\QBeta} \\[2mm]
		% \infrule{\G\V M:(\Pi x:\sigma.\tau)@A \andalso x\notin \text{FV}(M)}{\G\V(\lambda x:\sigma.M\ x)\E M: (\Pi x:\sigma.\tau)@A}{\QEta} \\[2mm]
		\infrule{\G\V M \E N : \tau@A}{\G\V \TBL_\alpha(\TB_\alpha M) \E N : \tau @A}{\QTBLTB} \\[2mm]
		\infrule{\G\V (\Lambda\alpha.M) : \forall\alpha.\tau@A}{\G\V (\Lambda\alpha.M)\ \epsilon \E M[\alpha \mapsto \epsilon] : \tau[\alpha \mapsto \epsilon]@A}{\QLambda} \\[2mm]
		\infrule{\G\V M:\tau@{A\alpha} \andalso \G\V M:\tau@A}{\G\V\%_\alpha M \E M : \tau@{A\alpha}}{\QPercent}
		\caption{Term Equivalence Rules}
		\label{fig:term-equivalence-rules}
	\end{center}
\end{figure}
